Police Arrests and Ethnic Background

### 2024-01-22 outputs/paper/paper.qmd
Additional Graphs
- Added Additional Graphs
- Fixed the styling
- Renamed the figures

#install.packages("tidyverse")
#install.packages("tinytex")
#install.packages("opendatatoronto")
# install.packages("kableExtra")
```

@@

library(tidyverse)
library(opendatatoronto)
library(tinytex)
library(knitr)
library(dplyr)
library(ggplot2)
library(janitor)
```

@@

```{r}
#| warning: false
#| message: false
#| include: false
# Adjusting and cleaning the name of the variable
cleaned_data <- data %>% 
  clean_names() %>% 
  clean_names(case = "snake")

# Isolating the columns that are needed for the analysis
cleaned_data <- cleaned_data |>
  select(arrest_year, arrest_month, perceived_race, age_group_at_arrest, youth_at_arrest_under_18_years, booked)

# Renaming the columns
cleaned_data <- cleaned_data |>
  rename(youth = youth_at_arrest_under_18_years, age_group = age_group_at_arrest)

# Saving the cleaned data
write.csv(cleaned_data, "../data/cleaned_data.csv")
```

```{r}
#| warning: false
#| message: false
#| include: false
# Make a table of the data with the number of arrests and strip searches by ethinic and identity based groups for the year 2020, and 2021
cleaned_data <- read.csv("../data/cleaned_data.csv")

@@

@@

@@

ggplot(
  table1, aes(x = `perceived_race`, y = `2020`, fill = `2021`)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Number of arrests by ethinic background for the year 2020, and 2021",
       x = "Ethinic and Identity Based Groups",
       y = "Number of Arrests") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_fill_gradient(low = "blue", high = "red") +
  theme(plot.title = element_text(hjust = 0.5))

@@

```{r}

>>>>

#install.packages("tidyverse")
#install.packages("tinytex")
#install.packages("opendatatoronto")
# install.packages("kableExtra")
# install.packages("patchwork")
```

@@

library(tidyverse)
library(opendatatoronto)
library(tinytex)
library(knitr)
library(dplyr)
library(ggplot2)
library(janitor)
library(patchwork)
```

@@

# Adjusting and cleaning the name of the variable
cleaned_data <- data %>% 
  clean_names() %>% 
  clean_names(case = "snake")

cleaned_data <- cleaned_data |>
  select(
    arrest_year, perceived_race, age_group_at_arrest, youth_at_arrest_under_18_years, booked, search_reason_cause_injury, search_reason_assist_escape, search_reason_possess_weapons, search_reason_possess_evidence)

# Combining the columns that are similar
cleaned_data <- cleaned_data %>%
  mutate(search_reason = case_when(
    search_reason_cause_injury == 1 ~ "1",
    search_reason_assist_escape == 1 ~ "1",
    search_reason_possess_weapons == 1 ~ "1",
    search_reason_possess_evidence == 1 ~ "1",
    TRUE ~ "0"
  ))

# Rename the columns
cleaned_data <- cleaned_data %>%
  rename(
    age_group = age_group_at_arrest,
    youth_at_arrest_under_18_years = youth_at_arrest_under_18_years
  )

@@

# Mutating for booked with NA values
cleaned_data <- cleaned_data %>%
  mutate(booked = case_when(
    is.na(booked) ~ "0",
    booked == 1 ~ "1",
    TRUE ~ "Not Booked"
  ))

cleaned_data <- cleaned_data |>
  select(
    arrest_year, perceived_race, age_group, youth_at_arrest_under_18_years, booked, search_reason)

write.csv(cleaned_data, "../data/cleaned_data.csv")

```{r}
#| warning: false
#| message: false
#| include: false
# Make a table of the data with the number of arrests and strip searches by ethinic and identity based groups for the year 2020, and 2021
# cleaned_data <- read.csv("../data/cleaned_data.csv")

@@

```{r}
cleaned_data <- read.csv("../data/cleaned_data.csv")

table <- cleaned_data |>
  group_by(arrest_year, perceived_race, search_reason) %>%
  summarise(n = n(), .groups = "drop") %>%
  arrange(desc(perceived_race)) %>%
  pivot_wider(names_from = arrest_year, values_from = n, values_fill = 0)

table <- pivot_longer(table, cols = c("2020", "2021"), names_to = "year", values_to = "value")

ggplot(table, aes(x = perceived_race, y = value, fill = factor(search_reason))) +
  geom_bar(stat = "identity", position = "fill") +
  facet_wrap(~ year, scales = "free_y") +
  labs(title = "Percent Stacked Bar Chart Based on 'booked'",
       x = "Perceived Race",
       y = "Percentage",
       fill = "Search Reason") +
  theme_minimal() +
  scale_fill_discrete(labels = c("Not Searched", "Searched")) +
  theme(legend.position = "right") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

@@

table1_long <- pivot_longer(table1, cols = c("2020", "2021"), names_to = "year", values_to = "value")
```

```{r}
#| warning: false
#| message: false
#| echo: false
table2 <- cleaned_data |>
  filter(booked == 1) |>
  group_by(arrest_year, perceived_race) |>
  summarise(n = n(), .groups = "drop") |>
  arrange(desc(n)) |>
  pivot_wider(names_from = arrest_year, values_from = n, values_fill = 0)

kable(table2,
      caption = "Arrests Booked by ethnic background for the year 2020, and 2021",
      align = "c",
      col.names = c("Ethinic and Identity Based Groups", "2020", "2021"),
      booktabs = TRUE)

table2_long <- pivot_longer(table2, cols = c("2020", "2021"), names_to = "year", values_to = "value")
```

```{r}
#| warning: false
#| message: false
#| echo: false
cleaned_data <- read.csv("../data/cleaned_data.csv")

cleaned_data <- cleaned_data |>
  select(
    arrest_year,
    perceived_race,
    booked
  )

# Group 
cleaned_data <- cleaned_data %>%
  group_by(arrest_year, perceived_race, booked) %>%
  summarise(n = n(), .groups = "drop") %>%
  arrange(desc(perceived_race)) %>%
  pivot_wider(names_from = arrest_year, values_from = n, values_fill = 0)

cleaned_data <- pivot_longer(cleaned_data, cols = c("2020", "2021"), names_to = "year", values_to = "value")
```

```{r police_arrests, fig.cap = "Number of arrests by ethinic background for the year 2020, and 2021", fig.width=10, fig.height=6, fig.align='center'}
ggplot(cleaned_data, aes(x = perceived_race, y = value, fill = factor(booked))) +
  geom_bar(stat = "identity") +
  facet_wrap(~ year, scales = "free_y") +
  labs(title = "Stacked Bar Chart Based on 'booked'",
       x = "Perceived Race",
       y = "Value",
       fill = "Booked") +
  theme_minimal() +
  scale_fill_discrete(labels = c("Not Booked", "Booked")) +
  theme(legend.position = "right") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

@@

```{r}
#| warning: false
#| message: false
#| echo: false
cleaned_data <- read.csv("../data/cleaned_data.csv")
cleaned_data <- cleaned_data |>
  select(
    arrest_year,
    perceived_race,
    age_group
  )

# Group 
cleaned_data <- cleaned_data %>%
  group_by(arrest_year, perceived_race, age_group) %>%
  summarise(n = n(), .groups = "drop") %>%
  arrange(desc(perceived_race)) %>%
  pivot_wider(names_from = arrest_year, values_from = n, values_fill = 0)

cleaned_data <- pivot_longer(cleaned_data, cols = c("2020", "2021"), names_to = "year"  ,values_to = "value")
```

```{r fig2, fig.cap = "Number of arrests by ethinic background for the year 2020, and 2021", fig.width=10, fig.height=6, fig.align='center'}
#| warning: false
#| message: false
#| echo: false
#| eval: true
# Dot Plot
ggplot(cleaned_data, aes(x = perceived_race, y = value, fill = factor(age_group))) +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 0.7, binwidth = 80) +
  facet_wrap(~ year, scales = "free_y") +
  labs(title = "Stacked Bar Chart Based on 'booked'",
       x = "Perceived Race",
       y = "Value",
       fill = "Booked") +
  theme_minimal() +
  theme(legend.position = "bottom") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```{r fig3, fig.cap = "Number of arrests by ethinic background for the year 2020, and 2021", fig.width=10, fig.height=6, fig.align='center'}
#| warning: false
#| message: false
#| echo: false
#| eval: true
# Bar plot
ggplot(cleaned_data, aes(x = perceived_race, y = value, fill = factor(age_group))) +
  geom_bar(stat = "identity") +
  facet_wrap(~ year, scales = "free_y") +
  labs(title = "Stacked Bar Chart Based on 'booked'",
       x = "Perceived Race",
       y = "Value",
       fill = "Booked") +
  theme_minimal() +
  theme(legend.position = "right") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
